---
title: "Prediction using Unsupervised ML"
author: "Nikhil Badri Prasad"
date: "11/21/2020"
output: word_document
---

Install and Load all the required libraries:
```{r}
#install.packages("tidyverse")
library(tidyverse)
library(cluster)
library(reshape2)
```

Import the iris dataset with the given link:
```{r}
data = iris
dim(data)
head(data)
```

Visualizing the data set:
Sepal-Length vs. Sepal-Width
```{r}
ggplot(data)+geom_point(aes(x = Sepal.Length, y = Sepal.Width), stroke = 2)+facet_wrap(~ Species)+ 
labs(x = "Sepal Length", y = "Sepal Width")+theme_bw()
```

Petal-Length vs. Petal-Width
```{r}
ggplot(data)+geom_point(aes(x = Petal.Length, y = Petal.Width), stroke = 2)+facet_wrap(~ Species)+ labs(x = "Petal Length", y = "Petal Width")+theme_bw()
```

Sepal-Length vs. Petal-Length
```{r}
ggplot(data)+geom_point(aes(x = Sepal.Length, y = Petal.Length), stroke = 2)+facet_wrap(~ Species)+ 
labs(x = "Sepal Length", y = "Petal Length")+theme_bw()
```

Sepal-Width vs. Pedal-Width
```{r}
ggplot(data)+geom_point(aes(x = Sepal.Width, y = Petal.Width), stroke = 2)+facet_wrap(~ Species)+ 
labs(x = "Sepal Width", y = "Pedal Width")+theme_bw()
```

Box-plots:
```{r}
ggplot(data)+
 geom_boxplot(aes(x = Species, y = Sepal.Length, fill = Species))+
 theme_bw()
ggplot(data)+
 geom_boxplot(aes(x = Species, y = Sepal.Width, fill = Species))+
 theme_bw()
ggplot(data)+
 geom_boxplot(aes(x = Species, y = Petal.Length, fill = Species))+
 theme_bw()
ggplot(data)+
 geom_boxplot(aes(x = Species, y = Petal.Width, fill = Species))+
 theme_bw()
```

k-means Clustering:
k-means clustering is a method of vector quantization, that aims to partition n observations into k clusters in which each observation belongs to the cluster with the nearest mean (cluster centers or cluster centroid), serving as a prototype of the cluster

Find the optimal number of clusters by Elbow Method:

```{r}
set.seed(123) # for reproduction
wcss <- vector()
for (i in 1:10) wcss[i] <- sum(kmeans(data[, -5], i)$withinss)
plot(1:10,
 wcss,
 type = 'b',
 main = paste("The Elbow Method"),
 xlab = "Number of Clusters",
 ylab = 'WCSS'
)
```

The elbow point : k(centers) = 3

Apply kmeans function to the feature columns:
```{r}
set.seed(123)
km <- kmeans( x = data[, -5] , centers = 3)
yclus <- km$cluster
table(yclus)
```

the kmeans has grouped the data into three clusters- 1, 2 & 3 having 50, 62 & 38 observations respectively.

Visualize the kmeans clusters:
```{r}
clusplot(data[, -5],
 yclus,
 lines = 0,
 shade = TRUE,
 color = TRUE,
 labels = 0,
 plotchar = FALSE,
 span = TRUE,
 main = paste("Clusters of Iris Flowers")
)
```

Compare the clusters:
```{r}
data$cluster.kmean <- yclus
cm <- table(data$Species, data$cluster.kmean)
cm
```

[(50 + 48 + 36)/150] = 89% of the k-means cluster output matched with the actual Species clusters. versicolor(Cluster 2) & virginica(Cluster 3) have some overlapping features which is also apparent from the cluster visualizations.